The SMT Competition 2015--2018

### 2018-10-12 2018/report/SMT-COMP.tex
Mainly SMT-LIB (work in progress)

\usepackage{pifont}
\usepackage{comment}

@@

The International Satisfiability Modulo Theories Competition
(SMT-COMP) is an annual competition between SMT solvers.  It was
instituted in~2005~\cite{TODO}, and is affiliated with the
International Workshop on Satisfiability Modulo Theories.  Solvers are
submitted to the competition by their developers, and pitted against
each other in a number of tracks and divisions.  The SMT Competition
was part of the Federated Logic Conference (FLoC) Olympic Games, which
comprised 14 competitions in various areas of computational logic, in
Vienna in~2014, and again in Oxford~(UK) in~2018.  It was last
described in a 2014 competition
report~\cite{DBLP:journals/jsat/CokDW14}.  In this paper, we report on
the design and selected results of the competition during the last
FLoC Olympiad, from 2015 to~2018.

@@

The original goals of the competition were to spur adoption of the
common, community-designed SMT-LIB formats (see
Section~\ref{sec:smtlib}), and to spark further advances in SMT,
especially for verification~\cite{smtcomp2005}.  These, together with
providing a useful yardstick of performance for users and developers
of SMT solvers, are still its main goals to date.  The competition has
been successful in establishing SMT-LIB as the de~facto standard
language for SMT solvers.  In recent years, the focus in this regard
has shifted towards promoting some of the newer extensions of the
SMT-LIB language~\cite{BarFT-RR-17}, such as its floating-point theory
or algebraic datatypes.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair.  This appointment
should happen ``within 30 days from the end of the current edition of
the SMT workshop''~\cite{smt-bylaws} but in practice has often
happened later.  The competition chair then assembles a team of
organizers, and distributes the work for the next edition of the
competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
different conference each year: with CAV in 2015, with IJCAR in 2016,
again with CAV in 2017, and with FLoC in~2018.  Consequently, its date
varies by several weeks.  Competition results were presented on
July~19 in 2015, on July~2 in 2016, on July~23 in 2017, and on July~13
in~2018.

@@

Around the initial solver deadline, the organizers also aim to publish
the latest version of the competition tools, and---in collaboration
with the SMT-LIB maintainers---to release a new version of the SMT
Library.  SMT-LIB releases were made on June~1 in 2015, on May~23 in
2016, on June~5 in 2017, and on May~20 in 2018.  Anyone may submit new
benchmarks to the SMT Library, and competitors are permitted to tune
their solvers accordingly.  Therefore it is important that all
benchmarks eligible for the competition are released at least some
time before the final solver submission deadline.  To give the SMT-LIB
maintainers sufficient time for curation, the deadline for new
benchmark contributions has usually been about four to six weeks
before the initial solver deadline, typically in April or early May.

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition web site
usually around mid-April.  The competition chair has ultimate
responsibility for the rules, but changes are often preceded by
discussions among the organizers or within the SMT community.  Such
discussions are initiated by a Call for Comments that is sent out to
the SMT-COMP mailing list~\cite{smtcomp-mailinglist} typically in
February or early March.  Publicly, this call is the first harbinger
of a new edition of the SMT Competition.

@@

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec (see Section~\ref{sec:starexec}), the cluster on which the
competition runs.

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based language that defines the
syntax and semantics of SMT solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows a possible representation of the
example formula from Section~\ref{sec:introduction} as a benchmark in
SMT-LIB syntax.

\begin{figure}
\begin{verbatim}
foo
\end{verbatim}
\caption{A benchmark problem in SMT-LIB syntax.}
\label{fig:smtlib-example}
\end{figure}

>>>>

\usepackage{pifont}
\usepackage{comment}
\usepackage{multirow}

@@

The International Satisfiability Modulo Theories Competition
(SMT-COMP) is an annual competition between SMT solvers.  It was
instituted in~2005~\cite{BdMS05}, and is affiliated with the
International Workshop on Satisfiability Modulo Theories.  Solvers are
submitted to the competition by their developers, and pitted against
each other in a number of tracks and divisions.  The SMT Competition
was part of the Federated Logic Conference (FLoC) Olympic Games, which
comprised 14 competitions in various areas of computational logic, in
Vienna in~2014, and again in Oxford~(UK) in~2018.  It was last
described in a 2014 competition
report~\cite{DBLP:journals/jsat/CokDW14}.  In this paper, we report on
the design and selected results of the competition during the last
FLoC Olympiad, from 2015 to~2018.

@@

The original goals of the competition were to spur adoption of the
common, community-designed SMT-LIB format (see
Section~\ref{sec:smtlib}), and to spark further advances in SMT,
especially for verification~\cite{BdMS05}.  These, together with
providing a useful yardstick of performance for users and developers
of SMT solvers, are still its main goals to date.  The competition has
been successful in establishing SMT-LIB as the de~facto standard
language for SMT solvers.  In recent years, the focus in this regard
has shifted towards promoting some of the newer extensions of the
SMT-LIB language, such as its floating-point theory or algebraic
datatypes~\cite{BarFT-RR-17}.

The competition is organized under the direction of the SMT Steering
Committee, which appoints the competition chair.  (This appointment
should happen ``within 30 days from the end of the current edition of
the SMT workshop''~\cite{smt-bylaws} but in practice has often
happened later.)  The competition chair then assembles a team of
organizers, and distributes the work for the next edition of the
competition.
%
The SMT Workshop includes a block of time to present the competitors
and results of the competition.  The workshop is affiliated with a
different conference each year: with CAV in 2015, with IJCAR in 2016,
again with CAV in 2017, and with FLoC in~2018.  Consequently, its date
varies by several weeks.  Competition results were presented on
July~19 in 2015, on July~2 in 2016, on July~23 in 2017, and on July~13
in~2018.

@@

Around the initial solver deadline, the organizers also aim to publish
the latest version of the competition tools, and---in collaboration
with the SMT-LIB maintainers---to release a new version of the SMT
Library (see Section~\ref{sec:smtlib}).  SMT-LIB releases were made on
June~1 in 2015, on May~23 in 2016, on June~5 in 2017, and on May~20 in
2018.  Anyone may submit new benchmarks to the SMT Library, and
competitors are permitted to tune their solvers accordingly.
Therefore all benchmarks eligible for the competition must be released
at least some time before the final solver submission deadline.  To
give the SMT-LIB maintainers sufficient time for curation, the
deadline for new benchmark contributions has usually been about four
to six weeks before the initial solver deadline, typically in April or
early May.

The competition rules are revised each year, and a (near final) draft
version of the rules is made public on the competition web site around
mid-April.  The competition chair has ultimate responsibility for the
rules, but changes are often preceded by discussions among the
organizers or within the SMT community.  Such discussions are
initiated by a Call for Comments that is sent out to the SMT-COMP
mailing list~\cite{smtcomp-mailinglist} typically in February or early
March.  Publicly, this call is the first harbinger of a new edition of
the SMT Competition.

@@

All problems used in the SMT Competition come from the SMT Library
(SMT-LIB).  This is a large collection of benchmark problems from
various sources.  The SMT Library is available from a public
repository~\cite{smtlib-repository}, and each new release is mirrored
on StarExec (see Section~\ref{sec:starexec}), the cluster on which the
competition runs.  Users of SMT solvers are encouraged to submit new
and interesting benchmarks to the library, which has grown from 1,352
benchmarks in 2005~\cite{BdMS05} to 347,011 benchmarks in 2018.

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based format that defines the
syntax and semantics of solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows the example formula from
Section~\ref{sec:introduction} as a benchmark in SMT-LIB syntax.

\begin{figure}
\begin{verbatim}
(set-info :smt-lib-version 2.6)
(set-logic QF_UFLIA)
(set-info :status unsat)
(declare-fun x () Int)
(declare-fun y () Int)
(declare-fun f (Int) Int)
(declare-fun P (Int) Bool)
(assert (and (<= x y) (<= y x) (P (- (f x) (f y))) (not (P 0))))
(check-sat)
(exit)
\end{verbatim}
\caption{A benchmark problem in SMT-LIB syntax.}
\label{fig:smtlib-example}
\end{figure}

The SMT Library classifies benchmarks according to their logic, that
is, according to the specific combination of background theories that
the benchmark uses.  Logic names are composed of letter groups that
refer to these background theories: {\tt A} or {\tt AX} for the theory
of arrays, {\tt BV} for the theory of bit-vectors, {\tt DT} for
algebraic datatypes, {\tt FP} for the floating-point theory, {\tt IA}
for integer arithmetic, {\tt RA} for real arithmetic, {\tt IRA} for
mixed integer and real arithmetic, {\tt S} for the theory of strings,
{\tt UF} for uninterpreted functions (i.e., free symbols).
Additionally, logics may impose further syntactic restrictions, such
as the absence of quantifiers ({\tt QF}), the restriction to
difference logic over the integers ({\tt IDL}) or reals ({\tt RDL}),
or to the (non-)linear fragment of arithmetic ({\tt L} or {\tt N},
respectively, before {\tt IA}, {\tt RA} or {\tt IRA}).  For instance,
the logic {\tt QF\_UFLIA} contains quantifier-free formulas with
uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

Additionally, the SMT Library distinguishes between incremental and
non-incremental benchmarks.  Non-incremental benchmarks contain a
single {\tt check-sat} command, which instructs the SMT solver to
determine the satisfiability of the benchmark.  (Valid solver
responses are {\tt sat}, {\tt unsat}, and {\tt unknown}.)  Incremental
benchmarks exercise additional solver features, such as the ability to
retract assertions, and typically contain multiple {\tt check-sat}
commands.  These benchmarks originate from applications such as
bounded model checking~\cite{TODO}, where SMT solvers interact with
other tools in a feedback loop.  The 2018 release of the SMT Library
contained 336,844 non-incremental benchmarks in 51 logics, as well
as~10,167 incremental benchmarks (with a total of nearly 32 million
{\tt check-sat} commands) in 26 logics.  Table~\ref{table:smtlib}
presents further data on the size of the SMT Library since~2015.
While the library usually grows from one year to the next due to the
inclusion of new benchmarks and logics, continuous curation efforts
may also cause benchmarks to be removed or reclassified on occasion,
thereby causing individual logics or even the entire library to
shrink.

\begin{table}
  \caption{Number of benchmarks and logics in the SMT Library}
  \label{table:smtlib}
  \centering
  \begin{tabular}{|r@{\ \ }l|r|r|r|r|}
    \hline
                               & & \multicolumn{1}{c|}{2015} & \multicolumn{1}{c|}{2016} & \multicolumn{1}{c|}{2017} & \multicolumn{1}{c|}{2018} \\
    \hline
    \multirow{2}{*}{Non-incremental $\{$} & benchmarks & 196,375 & 196,114 & 258,079 & 336,844 \\
                                          & logics     &      40 &      40 &      49 &      51 \\
    \multirow{2}{*}{Incremental $\{$}     & benchmarks &  10,019 &  10,024 &   6,262 &  10,167 \\
                                          & logics     &      15 &      15 &      22 &      26 \\
    \hline
  \end{tabular}  
\end{table}

Benchmarks in the SMT Library are annotated with additional
information, such as the source of the benchmark and its status (i.e.,
whether the benchmark is satisfiable or unsatisfiable) if this is
known.  (For incremental benchmarks, each {\tt check-sat} command has
a separate status annotation.)  To prevent solvers from simply looking
up the benchmark's status in the SMT Library, benchmarks in the
competition are lightly scrambled before they are presented to
solvers~\cite{DBLP:conf/cade/Weber16}.
